---
title: "Developing Data Products\nWeek 3 Assignment"
author: "Yang Zhang"
date: "8/26/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Instruction

```{r prepare_data, warning = FALSE, message = FALSE}
library(plotly)

diamonds_subset <- diamonds[sample(1: nrow(diamonds), 3000), ]
```

Create a web page presentation using R Markdown that features a plot created with Plotly. Host your webpage on either GitHub Pages, RPubs, or NeoCities. Your webpage must contain the date that you created the document, and it must contain a plot created with Plotly. 

This assignment uses a subset of the `diamonds` data set to:

* Show a plain scatterplot created by plotly
* Show how plotly could enhance the graph generated by ggplot2

## Plotly Graph

```{r plotly, warning = FALSE, message = FALSE}
p1 <- plot_ly(data = diamonds_subset,
              x = ~carat,
              y = ~price,
              color = ~cut,
              text = ~paste("Price: $", price ,
                            "<br>Carat: ", carat,
                            "<br>Cut: ", cut,
                            "<br>Color: ", color, 
                            "<br>Clarity:", clarity)) %>%
  layout(title = "Diamonds Price Distribution",
         xaxis = list(title = "Carat"),
         yaxis = list(title = "Price ($)"))

p1
```

## Enhanced ggplot2 Chart

```{r ggplot, warning = FALSE, message = FALSE}
p2 <- ggplot(data = diamonds_subset, aes(x = carat, y = price)) +
  geom_point(aes(color = color,
                 text = paste("clarity:", clarity))) +
  facet_wrap(~ cut) +
  labs(title = "Diamonds Price Distribution",
       xlab = "Carat",
       ylab = "Price ($)") +
  theme(legend.position = "none")

ggplotly(p2)
```